---
title: "Applying `riattrition` to Bowles et al. (2025)"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
execute:
  echo: true
  warning: false
  message: false
---

## Purpose

This note applies the updated `riattrition` package to:

- Bowles, Jeremy, Kevin Croke, Horacio Larreguy, Shelley Liu, and John Marshall. 2025. "Sustaining Exposure to Fact-Checks: Misinformation Discernment, Media Consumption, and Its Political Implications." *American Political Science Review* 119(4): 1864-1887.


## File Locations

```{r}
paper_pdf <- "/Users/yinghuizhou/Zotero/storage/K4VD7L2C/Bowles et al. - 2025 - Sustaining Exposure to Fact-Checks Misinformation Discernment, Media Consumption, and Its Political.pdf"
replication_dir <- "/Users/yinghuizhou/Library/CloudStorage/Dropbox/Projects/intro_rand_inf_missing/0_paper and data survey/3_apsr_2016_2025/APSR 2021-2025/dataset_apsr_2021-2026/raw_downloads/row0018"
package_dir <- "/Users/yinghuizhou/Library/CloudStorage/Dropbox/Projects/attrition application package/attrition application/riattrition"

list(
  paper_pdf = paper_pdf,
  replication_dir = replication_dir,
  package_dir = package_dir
)
```

## Summary of the Experiment

### Substantive question

The paper studies whether repeated exposure to fact-checks can make people more resistant to misinformation in a real-world setting, rather than only in a one-shot survey or lab environment.

The setting is South Africa, where WhatsApp is both widely used and highly relevant for the spread of political and public-health misinformation. The authors partnered with Africa Check and studied whether repeatedly sending fact-check content through WhatsApp changed participants' later ability to identify false information, their trust in media, their media behavior, and some downstream political and COVID-related attitudes.

### Recruitment and sample construction

The study recruited South African social media users through Facebook advertisements and then moved participants into a WhatsApp-based study workflow. Everyone first completed a baseline survey through a WhatsApp chatbot. According to the paper and replication materials:

- 11,672 people completed the baseline survey.
- 8,947 met the conditions needed to enroll in the experiment.
- 4,543 completed the endline survey after the six-month intervention period.

The paper also notes that enrolled participants were younger, more urban, more female, and more educated than the general population, but were closer to the relevant subgroup of Afrobarometer respondents who reported social media use.

### What exactly was randomized?

The core experimental intervention was repeated delivery of fact-check material over WhatsApp for six months.

Participants were assigned to:

- a pure control group that received no fact-checks; or
- one of four fact-check treatment arms that varied how the material was delivered.

The four main treatment formats were:

- a short text version;
- a short podcast;
- a long podcast; and
- an empathetic version of the long podcast.

All treated participants received the same underlying fact-check content every two weeks for six months, but the presentation format differed across arms.

### Additional cross-randomization

The paper also layered a second intervention on top of the main treatment assignment. Among treated participants, most were additionally assigned to receive monthly quizzes about recent fact-check content, with financial incentives for correct performance. Others, together with the control group, received placebo quizzes about unrelated content.

Substantively, it is a multi-arm intervention with an additional encouragement layer.

### Randomization process

The paper and online appendix describe randomization as being conducted separately within recruitment batches and within blocks of participants who were similar on pre-treatment characteristics. The blocking variables included demographics, social media use, trust in information sources, and misinformation knowledge.

The online appendix gives more detail on how those blocks were built:

- participants were enrolled every two weeks in rolling batches;
- within each batch, the authors used `blocktools`;
- blocks were formed using a greedy Mahalanobis-distance algorithm over seven predetermined baseline covariates;
- they first created blocks of size 38 so that treatment probabilities fit cleanly within block; and
- they then created nested sub-blocks of size 19 inside those larger blocks.

The appendix also notes that the paper's regression analyses use the larger block fixed effects of size 38, rather than the smaller sub-blocks of size 19, because attrition can leave some small sub-blocks without all treatment arms at endline.

So the design is:

- individual-level randomization;
- implemented within randomization blocks;
- repeated across rolling recruitment batches; and
- combined with a second randomization over quiz incentives.

This matters a lot for our application because the current `riattrition` package is designed for a binary treatment and complete randomization, while this paper's design is blocked and multi-arm.

### Treatment delivery

Treatment delivery happened through WhatsApp broadcast lists run through Africa Check. Treated participants were sent three fact-checks every two weeks for six months. Control participants received no fact-check delivery from Africa Check.

The fact-checks themselves mostly concerned false or misleading claims related to politics, public health, and social issues that had recently circulated in South Africa.

### Outcome measurement

At endline, the study measured:

- engagement with fact-check content;
- knowledge about treated fact-checks;
- willingness to consume fact-checks in the future;
- discernment between true and false stories;
- skepticism toward conspiracy theories;
- knowledge of verification strategies;
- attention to social media content;
- trust in social media and other information sources;
- information consumption, verification, and sharing behavior; and
- COVID-related and political attitudes and behavior.

The paper's main analyses often aggregate related survey items into inverse-covariance-weighted indexes.


## Load the Updated Package Code

For this application, we directly source the updated package code from the synchronized Dropbox package folder.

```{r}
source(file.path(package_dir, "R", "0_function_sharp_null.R"))
source(file.path(package_dir, "R", "0_function_sharp_null_twostep.R"))
source(file.path(package_dir, "R", "1_result_output.R"))
source(file.path(package_dir, "R", "2_visualization.R"))
```

## Load Replication Data

```{r}
library(dplyr)
library(haven)
library(knitr)

survey_data <- read_dta(file.path(replication_dir, "Data", "survey_data.dta"))

dim(survey_data)
```

## Replication Data Orientation

The replication data already contains the variables we need for a first application:

- `t0`: pooled treatment assignment
- `t5`: disaggregated treatment assignment
- `block_id`: randomization block
- `in_endline`: whether the respondent completed the endline survey
- several endline outcomes beginning with `end_icw_`

```{r}
survey_data %>%
  transmute(
    pooled_arm = as_factor(t0),
    disagg_arm = as_factor(t5),
    in_endline = in_endline
  ) %>%
  summarise(
    n_total = n(),
    n_endline = sum(in_endline == 1, na.rm = TRUE),
    n_not_endline = sum(in_endline == 0, na.rm = TRUE)
  )
```

```{r}
table(as_factor(survey_data$t0), useNA = "ifany")
table(as_factor(survey_data$t5), useNA = "ifany")
length(unique(survey_data$block_id))
```

## Application Strategy

The published experiment is richer than the original package interface in two ways:

1. treatment assignment is multi-arm rather than binary;
2. randomization was blocked rather than complete-randomization-without-blocks.

With the updated wrapper, we can now handle the blocked-randomization feature by
supplying the block identifier to `ri_sharp_attrition()`. We still simplify the multi-arm
design into a binary contrast for the first application.

So for a first package application, we use this mapping:

- keep only the `Control` and `Pooled treatment` arms from `t0`;
- drop the `Placebo incentives` arm for the first application;
- treat the result as a binary-treatment experiment;
- preserve the original blocked randomization by passing `block_id`;
- use `missing = "general"` as the default missingness assumption; and
- treat this as an exploratory application intended to show where the package still needs to improve.

This is not the final or ideal analysis for the paper. It is a deliberate first-pass application of the current package.

### Construct a binary treatment dataset

```{r}
app_data <- survey_data %>%
  mutate(
    pooled_arm = as_factor(t0),
    Z = case_when(
      pooled_arm == "Control" ~ 0L,
      pooled_arm == "Pooled treatment" ~ 1L,
      TRUE ~ NA_integer_
    )
  ) %>%
  filter(!is.na(Z))

app_data %>%
  count(pooled_arm)
```

## Attrition Summary for the Binary Mapping

```{r}
attrition_summary <- app_data %>%
  group_by(pooled_arm) %>%
  summarise(
    n = n(),
    endline_completed = sum(in_endline == 1, na.rm = TRUE),
    endline_missing = sum(in_endline == 0, na.rm = TRUE),
    endline_completion_rate = mean(in_endline == 1, na.rm = TRUE)
  )

kable(attrition_summary, digits = 3)
```


## Analysis Setup


```{r}
nperm <- 1000
set.seed(11215)
```

We will use two outcomes:

- `end_icw_discernment`
- `end_icw_treat_knowledge`

For the paper comparison later on, these map to:

- `Table F4`, pooled estimation, `Pooled treatment` row for discernment;
- `Table F2`, pooled estimation, `Pooled treatment` row for treatment knowledge.

## Direct Wrapper Application with `ri_sharp_attrition()`

Before unpacking the mechanics, we first run the high-level wrapper.

```{r}
discernment_direct <- ri_sharp_attrition(
  Z = app_data$Z,
  Y = app_data$end_icw_discernment,
  c = 0,
  missing = "general",
  class = "RS",
  method.list = list(name = "Wilcoxon"),
  block = app_data$block_id,
  nperm = nperm,
  include_twostep = FALSE
)

treat_knowledge_direct <- ri_sharp_attrition(
  Z = app_data$Z,
  Y = app_data$end_icw_treat_knowledge,
  c = 0,
  missing = "general",
  class = "RS",
  method.list = list(name = "Wilcoxon"),
  block = app_data$block_id,
  nperm = nperm,
  include_twostep = FALSE
)
```

```{r}
discernment_direct
```

```{r}
treat_knowledge_direct
```

## Visualizing general versus complete-case inference

The new `plot_ri_test_stat()` helper uses the same sharp-null machinery as
`ri_sharp_attrition()`, but turns the output into a visual summary. In the
default histogram view:

- the bars show the empirical randomization null distribution;
- the dashed reference marks the observed test statistic; and
- the highlighted right tail corresponds to the p-value
  \eqn{\Pr(T_{null} \geq T_{obs})}.

For discernment, it is especially useful to compare:

- `analysis = "general"`, which treats attrition as a core missing-data problem;
- `analysis = "complete-case"`, which restricts attention to endline respondents only.

```{r, include=FALSE}
set.seed(11215)
discernment_general_plot <- plot_ri_test_stat(
  Z = app_data$Z,
  Y = app_data$end_icw_discernment,
  c = 0,
  missing = "general",
  class = "RS",
  method.list = list(name = "Wilcoxon"),
  block = app_data$block_id,
  nperm = nperm
)

set.seed(11215)
discernment_complete_case_plot <- plot_ri_test_stat(
  Z = app_data$Z,
  Y = app_data$end_icw_discernment,
  c = 0,
  analysis = "complete-case",
  class = "RS",
  method.list = list(name = "Wilcoxon"),
  block = app_data$block_id,
  nperm = nperm
)
```

### General missingness

```{r, fig.width=9, fig.height=5}
set.seed(11215)
plot_ri_test_stat(
  Z = app_data$Z,
  Y = app_data$end_icw_discernment,
  c = 0,
  missing = "general",
  class = "RS",
  method.list = list(name = "Wilcoxon"),
  block = app_data$block_id,
  nperm = nperm
)
```

### Complete-case analysis

```{r, fig.width=9, fig.height=5}
set.seed(11215)
plot_ri_test_stat(
  Z = app_data$Z,
  Y = app_data$end_icw_discernment,
  c = 0,
  analysis = "complete-case",
  class = "RS",
  method.list = list(name = "Wilcoxon"),
  block = app_data$block_id,
  nperm = nperm
)
```

The returned objects also store the plotted values, so we can compare the two
analyses directly:

```{r}
bind_rows(
  tibble(
    analysis = discernment_general_plot$mode_label,
    label = discernment_general_plot$assumption,
    test_statistic = discernment_general_plot$test_name,
    observed_statistic = discernment_general_plot$stat_obs,
    p_value = discernment_general_plot$p_value,
    design = discernment_general_plot$design,
    nperm = discernment_general_plot$nperm
  ),
  tibble(
    analysis = discernment_complete_case_plot$mode_label,
    label = discernment_complete_case_plot$assumption,
    test_statistic = discernment_complete_case_plot$test_name,
    observed_statistic = discernment_complete_case_plot$stat_obs,
    p_value = discernment_complete_case_plot$p_value,
    design = discernment_complete_case_plot$design,
    nperm = discernment_complete_case_plot$nperm
  )
) %>%
  kable(digits = 3)
```

- Under `analysis = "complete-case"`, the test drops missing outcomes and works
  only with the `4543` endline respondents. In this observed-only sample, the
  discernment outcome is higher in treatment than in control, and the observed
  Wilcoxon rank-sum statistic is about `1682.755`. The corresponding blocked
  complete-case null distribution is much farther left, roughly from `1657.510`
  to `1668.716`, so none of the `1000` null draws is at least as large as the
  observed statistic. That is why the empirical p-value is `0.000`.
- Under `analysis = "general"`, the test keeps all `8947` randomized units and
  applies the package's most conservative completion rule under general
  missingness. Here there are `4404` missing outcomes in total, including
  `3260` missing treated outcomes and `1144` missing control outcomes. Under
  the sharp null with `c = 0`, the missing treated outcomes are placed at
  `-Inf`, while the missing control outcomes are placed at `Inf`. This pushes
  the observed rank-sum statistic sharply downward to about `2665.175`, whereas
  the blocked general-missingness null distribution is much farther right,
  roughly from `3289.456` to `3303.063`. Every one of the `1000` null draws is
  at least as large as the observed statistic, so the empirical p-value is
  `1.000`.

## Sensitivity Check: `missing = "mn"`

Because the treated attrition rate is slightly higher than the control
attrition rate in this application, `mn` is a natural directional sensitivity
assumption to inspect.

Here we rerun the discernment outcome under `missing = "mn"` so that we can
compare the printed output directly to the default `general` case.

```{r}
discernment_mn <- ri_sharp_attrition(
  Z = app_data$Z,
  Y = app_data$end_icw_discernment,
  c = 0,
  missing = "mn",
  class = "RS",
  method.list = list(name = "Wilcoxon"),
  block = app_data$block_id,
  block_missing_sum = TRUE,
  nperm = nperm,
  include_ci = TRUE,
  include_twostep = FALSE
)
```

```{r}
discernment_mn
```

Notice that `ri_sharp_attrition()` always reports the `general` row first. When
the user requests `missing = "mn"`, the table adds a second row for `mn`
underneath the default `general` result.

```{r}
mn_comparison <- bind_rows(
  discernment_direct$results %>%
    mutate(output_call = "general only"),
  discernment_mn$results %>%
    mutate(output_call = "missing = mn")
) %>%
  select(
    output_call,
    `Missingness assumption`,
    Method,
    `Test stat.`,
    `Test stat. value`,
    `P-value`,
    `CI lower`,
    `CI upper`
  )

kable(mn_comparison, digits = 3)
```

The rest of this note reconstructs the sharp-null result step by step using the
lower-level functions that `ri_sharp_attrition()` calls internally.

## Step-by-Step Walkthrough: Discernment

### Prepare inputs

```{r}
discernment_Y <- app_data$end_icw_discernment
discernment_Z <- app_data$Z
discernment_block <- app_data$block_id
```

### Compute missingness counts with `.ri_missing_counts()`

This function summarizes the total number of observations, missing outcomes,
and attrition rates by treatment status.

```{r}
discernment_counts <- .ri_missing_counts(
  Z = discernment_Z,
  Y = discernment_Y
)

discernment_counts_tbl <- tibble(
  statistic = c(
    "Number of Obs.",
    "Observed outcome",
    "Missing outcome",
    "Observed treated",
    "Observed control",
    "Missing treated",
    "Missing control",
    "Overall attrition rate",
    "Treated attrition rate",
    "Control attrition rate"
  ),
  value = c(
    discernment_counts$n_total,
    discernment_counts$n_obs,
    discernment_counts$n_miss,
    discernment_counts$n_obs_treat,
    discernment_counts$n_obs_control,
    discernment_counts$n_miss_treat,
    discernment_counts$n_miss_control,
    discernment_counts$attrition_rate_total,
    discernment_counts$attrition_rate_treat,
    discernment_counts$attrition_rate_control
  )
)

kable(discernment_counts_tbl, digits = 3)
```

### Compute block diagnostics with `.ri_block_diagnostics()`

Because this is a blocked experiment, we also summarize attrition patterns
within blocks.

```{r}
discernment_block_diag <- .ri_block_diagnostics(
  Z = discernment_Z,
  Y = discernment_Y,
  block = discernment_block
)

discernment_block_tbl <- tibble(
  statistic = c(
    "Number of blocks",
    "Mean within-block attrition diff",
    "Median within-block attrition diff"
  ),
  value = c(
    discernment_block_diag$n_blocks,
    discernment_block_diag$mean_within_block_attrition_diff,
    discernment_block_diag$median_within_block_attrition_diff
  )
)

kable(discernment_block_tbl, digits = 3)
```

### Generate blocked permutations with `.ri_default_assignments()`

For `missing = "general"` and `block != NULL`, `ri_sharp_attrition()` delegates to
`.ri_default_assignments()`, which in turn calls `.ri_assign_blocked()`.

The key idea is:

- treatment labels are permuted within each block;
- the number treated in each block is preserved; and
- no unit ever moves across blocks.

```{r}
set.seed(11215)
discernment_Z_perm <- .ri_default_assignments(
  Z = discernment_Z,
  Y = discernment_Y,
  missing = "general",
  block = discernment_block,
  nperm = nperm
)

dim(discernment_Z_perm)
```

Every permutation should preserve the total number treated:

```{r}
tibble(
  permutation = 1:5,
  treated_total = colSums(discernment_Z_perm[, 1:5, drop = FALSE])
) %>%
  kable()
```

And within selected blocks, the treated count should also be preserved:

```{r}
selected_blocks <- unique(discernment_block)[1:3]

block_perm_check <- bind_rows(lapply(selected_blocks, function(b) {
  idx <- which(discernment_block == b)
  tibble(
    block = as.character(b),
    observed_treated = sum(discernment_Z[idx]),
    perm_1 = sum(discernment_Z_perm[idx, 1]),
    perm_2 = sum(discernment_Z_perm[idx, 2]),
    perm_3 = sum(discernment_Z_perm[idx, 3]),
    perm_4 = sum(discernment_Z_perm[idx, 4]),
    perm_5 = sum(discernment_Z_perm[idx, 5])
  )
}))

kable(block_perm_check)
```

### Build the completed outcome under the sharp null

For `missing = "general"` and `c = 0`, `.ri_sharp_details()` completes the
outcome vector in the most conservative way:

- observed treated values stay as observed;
- missing treated values become `-Inf`;
- observed control values stay as observed; and
- missing control values become `Inf`.

This is the completed outcome vector that feeds the rank-based test statistic.

```{r}
discernment_M <- as.numeric(!is.na(discernment_Y))
discernment_Y0_imp <- discernment_Y - discernment_Z * 0
discernment_Y0_com_imp <- rep(NA_real_, length(discernment_Y))

discernment_Y0_com_imp[discernment_Z == 1 & discernment_M == 1] <- discernment_Y0_imp[discernment_Z == 1 & discernment_M == 1]
discernment_Y0_com_imp[discernment_Z == 1 & discernment_M == 0] <- -Inf
discernment_Y0_com_imp[discernment_Z == 0 & discernment_M == 1] <- discernment_Y0_imp[discernment_Z == 0 & discernment_M == 1]
discernment_Y0_com_imp[discernment_Z == 0 & discernment_M == 0] <- Inf

tibble(
  component = c(
    "Finite observed values",
    "Treated missing imputed to -Inf",
    "Control missing imputed to Inf"
  ),
  n = c(
    sum(is.finite(discernment_Y0_com_imp)),
    sum(discernment_Y0_com_imp == -Inf),
    sum(discernment_Y0_com_imp == Inf)
  )
) %>%
  kable()
```

### Compute the observed test statistic with `test_stat()`

With `class = "RS"` and `method.list = list(name = "Wilcoxon")`, the package
uses a Wilcoxon rank-sum test statistic.

```{r}
discernment_stat_obs <- test_stat(
  Z = discernment_Z,
  Y = discernment_Y0_com_imp,
  class = "RS",
  method.list = list(name = "Wilcoxon")
)

discernment_stat_obs
```

### Compute the randomization distribution with `null_dist()`

`null_dist()` uses the blocked permutation matrix `discernment_Z_perm` to build
the null distribution of the same rank-sum statistic.

```{r}
discernment_stat_null <- null_dist(
  n = length(discernment_Z),
  m = sum(discernment_Z),
  class = "RS",
  method.list = list(name = "Wilcoxon"),
  Z.perm = discernment_Z_perm,
  nperm = nperm
)

summary(discernment_stat_null)
```

### Compute the randomization p-value manually

For the one-sided sharp-null test implemented here, the p-value is the fraction
of null draws that are at least as large as the observed statistic.

```{r}
discernment_p_manual <- mean(discernment_stat_null >= discernment_stat_obs)

discernment_p_manual
```

### Verify the same result with `.ri_sharp_details()`

`.ri_sharp_details()` bundles together the same logic: missingness counts,
worst-case completion under the chosen assumption, the observed test statistic,
and the sharp-null p-value.

```{r}
discernment_sharp_details <- .ri_sharp_details(
  Z = discernment_Z,
  Y = discernment_Y,
  c = 0,
  missing = "general",
  class = "RS",
  method.list = list(name = "Wilcoxon"),
  stat.null = NULL,
  Z.perm = discernment_Z_perm,
  nperm = nperm
)

discernment_sharp_details
```

### Compare the manual result to `ri_sharp_attrition()`

Finally, we call `ri_sharp_attrition()` using the exact same permutation matrix. This lets
us check that the wrapper is simply presenting the same calculation in a richer
user-facing format.

```{r}
discernment_from_wrapper <- ri_sharp_attrition(
  Z = discernment_Z,
  Y = discernment_Y,
  c = 0,
  missing = "general",
  class = "RS",
  method.list = list(name = "Wilcoxon"),
  block = discernment_block,
  Z.perm = discernment_Z_perm,
  nperm = nperm,
  include_ci = FALSE,
  include_twostep = FALSE
)

discernment_comparison <- tibble(
  quantity = c("Observed test statistic", "P-value"),
  manual = c(discernment_stat_obs, discernment_p_manual),
  `.ri_sharp_details()` = c(discernment_sharp_details$stat_obs, discernment_sharp_details$p_value),
  `ri_sharp_attrition()` = c(
    discernment_from_wrapper$results$`Test stat.`[1],
    discernment_from_wrapper$results$`P-value`[1]
  )
)

kable(discernment_comparison, digits = 3)
```

```{r}
discernment_from_wrapper
```

The manual implementation, `.ri_sharp_details()`, and `ri_sharp_attrition()` agree exactly.

## Treatment Knowledge

The same function chain applies to `end_icw_treat_knowledge`. To keep the note
readable, we use one compact helper that runs the same sequence of functions and
stores the intermediate results.

```{r}
run_manual_general_rs <- function(Y, Z, block, outcome_name, nperm = 1000) {
  counts <- .ri_missing_counts(Z = Z, Y = Y)
  block_diag <- .ri_block_diagnostics(Z = Z, Y = Y, block = block)
  Z.perm <- .ri_default_assignments(
    Z = Z, Y = Y, missing = "general", block = block, nperm = nperm
  )

  M <- as.numeric(!is.na(Y))
  Y0.com.imp <- rep(NA_real_, length(Y))
  Y0.com.imp[Z == 1 & M == 1] <- Y[Z == 1 & M == 1]
  Y0.com.imp[Z == 1 & M == 0] <- -Inf
  Y0.com.imp[Z == 0 & M == 1] <- Y[Z == 0 & M == 1]
  Y0.com.imp[Z == 0 & M == 0] <- Inf

  stat_obs <- test_stat(
    Z = Z,
    Y = Y0.com.imp,
    class = "RS",
    method.list = list(name = "Wilcoxon")
  )

  stat_null <- null_dist(
    n = length(Z),
    m = sum(Z),
    class = "RS",
    method.list = list(name = "Wilcoxon"),
    Z.perm = Z.perm,
    nperm = nperm
  )

  p_value <- mean(stat_null >= stat_obs)

  sharp_details <- .ri_sharp_details(
    Z = Z,
    Y = Y,
    c = 0,
    missing = "general",
    class = "RS",
    method.list = list(name = "Wilcoxon"),
    stat.null = NULL,
    Z.perm = Z.perm,
    nperm = nperm
  )

  direct <- ri_sharp_attrition(
    Z = Z,
    Y = Y,
    c = 0,
    missing = "general",
    class = "RS",
    method.list = list(name = "Wilcoxon"),
    block = block,
    Z.perm = Z.perm,
    nperm = nperm,
    include_ci = FALSE,
    include_twostep = FALSE
  )

  observed_means <- tibble(
    outcome = outcome_name,
    mean_treated = mean(Y[Z == 1 & !is.na(Y)]),
    mean_control = mean(Y[Z == 0 & !is.na(Y)]),
    observed_diff = mean_treated - mean_control
  )

  comparison <- tibble(
    outcome = outcome_name,
    manual_stat = stat_obs,
    wrapper_stat = sharp_details$stat_obs,
    ri_test_stat = direct$results$`Test stat.`[1],
    manual_p = p_value,
    wrapper_p = sharp_details$p_value,
    ri_test_p = direct$results$`P-value`[1]
  )

  list(
    counts = counts,
    block_diag = block_diag,
    observed_means = observed_means,
    comparison = comparison,
    direct = direct
  )
}

treat_knowledge_walkthrough <- run_manual_general_rs(
  Y = app_data$end_icw_treat_knowledge,
  Z = app_data$Z,
  block = app_data$block_id,
  outcome_name = "end_icw_treat_knowledge",
  nperm = nperm
)
```

```{r}
kable(treat_knowledge_walkthrough$observed_means, digits = 3)
```

```{r}
kable(treat_knowledge_walkthrough$comparison, digits = 3)
```

```{r}
treat_knowledge_walkthrough$direct
```

Again, the step-by-step implementation and `ri_sharp_attrition()` match exactly.

## Compare to the Paper's Original Regression Tables

The paper's corresponding regression tables are in the online appendix:

- `Table F2: Treatment knowledge`
- `Table F4: Discernment`

For both tables, the relevant comparison for our first-pass binary recoding is:

- section `A. Pooled estimation`;
- row `Pooled treatment`;
- columns `(1)` and `(2)`.

Column `(1)` uses block fixed effects and no additional controls, while column
`(2)` adds LASSO-selected controls. Both are estimated on endline respondents.

```{r}
paper_comparison <- tibble(
  outcome = c("end_icw_treat_knowledge", "end_icw_discernment"),
  paper_table = c(
    "Table F2, pooled estimation, Pooled treatment, cols (1)-(2)",
    "Table F4, pooled estimation, Pooled treatment, cols (1)-(2)"
  ),
  paper_coef_col1 = c(0.411, 0.058),
  paper_coef_col2 = c(0.414, 0.063),
  paper_p_col1 = c(0.000, 0.048),
  paper_p_col2 = c(0.000, 0.034),
  paper_observations = c(4543, 4543),
  app_observed_diff = c(
    treat_knowledge_walkthrough$observed_means$observed_diff,
    mean(app_data$end_icw_discernment[app_data$Z == 1 & !is.na(app_data$end_icw_discernment)]) -
      mean(app_data$end_icw_discernment[app_data$Z == 0 & !is.na(app_data$end_icw_discernment)])
  ),
  app_general_p = c(
    treat_knowledge_walkthrough$direct$results$`P-value`[1],
    discernment_from_wrapper$results$`P-value`[1]
  ),
  app_randomized_n = c(nrow(app_data), nrow(app_data)),
  app_observed_outcome_n = c(
    sum(!is.na(app_data$end_icw_treat_knowledge)),
    sum(!is.na(app_data$end_icw_discernment))
  )
)

kable(paper_comparison, digits = 3)
```

### Interpretation of the comparison

For treatment knowledge:

- the paper's pooled treatment coefficient is `0.411` in column `(1)` and
  `0.414` in column `(2)` of `Table F2`;
- our observed complete-case treated-control difference in the simplified
  `Control` versus `Pooled treatment` sample is about `0.392`; and
- our `ri_sharp_attrition()` result under `missing = "general"` gives a sharp-null
  p-value of `1.000`.

For discernment:

- the paper's pooled treatment coefficient is `0.058` in column `(1)` and
  `0.063` in column `(2)` of `Table F4`;
- our observed complete-case treated-control difference is about `0.059`; and
- our `ri_sharp_attrition()` result under `missing = "general"` again gives a sharp-null
  p-value of `1.000`.

The coefficient comparison and the inference
comparison are answering different questions.

- The paper's regression tables condition on endline respondents, use all
  treatment arms, and estimate OLS models with block fixed effects.
- Our current package application keeps all randomized units in the binary
  `Control` versus `Pooled treatment` sample and treats attrition as missing
  potential outcomes under the user-chosen assumption.
- Under the default `general` assumption, the sharp-null test is extremely
  conservative because missing treated outcomes are sent to `-Inf` and missing
  control outcomes are sent to `Inf`.
